---
title: "In-Class Exercise 04"
author: "Ong Chae Hui"
date: 6 May 2023
date-modified: "`r Sys.Date()`"
execute:
  echo: true
  eval: true
  warning: false
---

## 1. Getting Started

### 1.1. Installing and Loading the required R Packages

In this exercise using Exam_data, we will be using **tidyverse**, [**rstatix**](https://cran.r-project.org/web/packages/rstatix/index.html), **gt** and **patchwork**.

```{r}
pacman::p_load(tidyverse, rstatix, gt, patchwork)
```

### 1.2. Importing Data (Exam_data)

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

### 1.3. Visualising Normal Distribution

A Q-Q plot (Quantile-Quantile plot) is used to assess whether a set of data points are normally distributed. 

if the data is normally distrbuted, the points in a Q-Q plot will lie on a straight diagonal line.
Conversely, if the points deviate significantly from the straight diagonal line, then it's less likely that the data is normally distributed.

::: panel-tabset
## The Plot
```{r}
#| echo: false
#| fig-width: 4
#| fig-height: 4
ggplot(exam_data,
       aes(sample=ENGLISH)) + 
  stat_qq() + 
  stat_qq_line()
```

## The Code Chunk
```{r}
#| eval: false
#| fig-width: 4
#| fig-height: 4

ggplot(exam_data,
       aes(sample=ENGLISH)) + 
  stat_qq() + 
  stat_qq_line()
```
:::

::: {.callout-note}
We can see that the points deviate significantly form the straight diagnoal line. This is a clear indication that the set of data is not normally distributed. 
:::


### 1.4. Runnig Shapiron Test

**png**, **webshot2** packages will be required to run the following codes.

::: panel-tabset

## The Plot
```{r}
#| echo: false

qq <- ggplot(exam_data,
             aes(sample=ENGLISH)) + 
  stat_qq() +
  stat_qq_line()

# running shapiro test and save into gt() format
sw_t <- exam_data %>%
  shapiro_test(ENGLISH) %>%
  gt()


# converting the sw_t into an image file (png)
tmp <- tempfile(fileext = '.png')
gtsave(sw_t, tmp)
table_png <- png::readPNG(tmp,
                          native = TRUE)

qq + table_png

```

## The Code Chunk

```{r}
#| eval: false

qq <- ggplot(exam_data,
             aes(sample=ENGLISH)) + 
  stat_qq() +
  stat_qq_line()

# running shapiro test and save into gt() format
sw_t <- exam_data %>%
  shapiro_test(ENGLISH) %>%
  gt()


# converting the sw_t into an image file (png)
tmp <- tempfile(fileext = '.png')
gtsave(sw_t, tmp)
table_png <- png::readPNG(tmp,
                          native = TRUE)

qq + table_png
```
:::










